Optimal Approximation of Discrete Random Variables \\ for Estimation of Probabilities of Missing Deadlines

### 2018-09-05 Kolmogorov_Approx.tex
Updates from ShareLaTeX

\DeclareMathOperator{\supp}{support}
\DeclareMathOperator{\support}{support}
\DeclareMathOperator{\Trim}{Trim}
\DeclareMathOperator{\LTrim}{LTrim}
\SetKwFunction{mEqOne}{mEqOne} 
\DeclareMathOperator{\KlmApprox}{KolmogorovApprox}
%\DeclareMathOperator{\OptTrim}{KolmogorovApprox}
\DeclareMathOperator{\OptTrim}{OptTrim}

@@

@@

The size of a random variable is measured here by the size of its support, the set of possible outcomes, $m=|\support(X)|=|\{x\colon Pr(X=x) \neq 0\}|$. When probability mass functions are maintained as explicit tables, as done in many implementations of statistical software, the size of the support of the variable is proportional to the amount of memory needed to store it and to the complexity of the computations that manipulate it.

@@

In the technical level, the problem we study in this paper is similar to the problem of approximating a set of 2-D points by a step function. The study of this problem was motivated by query optimisation and histogram constructions in database management systems~\cite{applf12,applf13,applf14,applf17,applf18,Fournier2011}. There are, however, two technically significant differences between the problem studied in the context of databases and the problem we analyse in this paper. The first difference is that in the context of approximation of random variables, the step function (which is the cumulative distribution function in our context)  must end with a value one, since we are dealing with random variables which sums to one. The second difference is that the first step is not counted because there is no need to put a value in the support of the approximated random variable to generate this first step. These cannot be addressed by adding a constant (two) to $m$ because the first step is always present and because the requirement to end with the value one, restricts the set of eligible step functions. See Section~\ref{sec:alg} for more details.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An algorithm for optimal approximation}\label{sec:algEfficient}

@@

\begin{theorem}\label{the:complexityDual}
	The $dual(X,\varepsilon)$ algorithm runs in time $O(n)$, using $O(n)$ memory where $n=|\support(X)|$.
\end{theorem}
\begin{proof}
	This algorithm describes one pass over the random variable $X$. Lines 3-4 and lines 5-6 are easy to follow, each takes $O(n)$ in the worst case. Lines 8-12 also describe one pass over $X$ since the counter $e$ is updated to $e+1$ at most $n$ times. All together we get run-time complexity of $O(3n)$ = $O(n)$. We are constructing the set $S$ which is of size $n$ in the worst case, therefore, memory complexity is also $O(n)$. 
\end{proof}

\begin{algorithm}
	\DontPrintSemicolon
	Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $Pr(X \leq x_i)=c_i$ and $\supp(X)=\{x_1 < \cdots < x_n\}$.\;
    $E \gets \emptyset$\; 
    \For{$i \gets 1$ \textbf{to} $n-1$}{
    	\For{$j \gets i+1$ \textbf{to} $n-1$}{
        		\If{$i=1$} {
                	$E \gets E \cup \{ c_j \}$
                }
        		$E \gets E \cup \{ (c_j-c_i)/2 \}$
        }
        $E \gets E \cup \{ 1 - c_i \}$
    }
    Let $e_1<\cdots<e_{n'}$ be such that $E=\{e_1,\dots,e_{n'}\}$ \;

@@

	\caption{$approx(X,m)$}   
	\label{alg:naive}
\end{algorithm}

\begin{theorem}\label{the:complexityDual}
	The $approx(X,m)$ algorithm runs in time $O(n^2log(n))$, using $O(n^2)$ memory where $n=|\support(X)|$.
\end{theorem}
\begin{proof}
	In the first part of the algorithm, lines 2-8, we construct the set $E$ which takes $O(n^2)$ run-time. In the second part of the algorithm, line 9, we sort the set $E$ which takes $O(n^2log(n^2)) = O(n^2log(n))$  run-time. The third part of the algorithm, lines 10-19, describes a binary sort over the set $E$ where in each step of the sorting we run the $dual$ algorithm, which takes $O(n log(n^2)) = O(n log(n))$ run-time. We run this part twice then we get $O(2n log(n))$. All together, the run-time complexity is $O(n^2+n^2log(n)+2n log(n)) = O(n^2log(n))$ and the memory complexity is for set $E$ which is $n^2$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An algorithm for optimal approximation}\label{sec:alg}
In the scope of this section, let $X$ be a given random variable with a finite support of size $n$, and let  $0<m\leq n$ be a given complexity bound. The section evolves by developing notations and by collecting facts towards an algorithm for finding an optimal $m$-approximation of $X$.

The first useful fact is that it is enough to limit our search to approximations $X'$s such that $\support(X') \subseteq \support(X)$:

%Next, note that every random variable $X''$ with support of size at most $m$ that is contained in $\support(X)$ can be described by first setting the (at most $m$) elements of the support of $X''$; then for every such option, determine $X''$ by setting probability values for the elements in the chosen support of $X'$, and setting $0$ for rest of the elements.

For a set $S\subseteq \support(X)$, let $\mathbb{X}_S$ denote the set of random variables whose supports are contained in $S$. In Step 1 below, we identify a random variable in $\mathbb{X}_S$  that minimizes the Kolmogorov distance from $X$. We denote the Kolmogorov distance between this variable and $X$ by $\varepsilon(X,S)$. Then, in Step 2, we show how to efficiently find a set $S \subseteq \support(X)$ whose size is smaller or equal to $m$ that minimizes $\varepsilon(X,S)$. Then, in Step 3, an optimal $m$-approximation is constructed by taking a minimal approximation in $\mathbb{X}_S$ where $S$ is the set of size $m$ that that minimizes $\varepsilon(X,S)$.

\subsection*{Step 1: Finding an $X'$ in $\mathbb{X}_S$ that minimizes $d_K(X,X')$}

We first fix a set $S\subseteq \support(X)$ and among all the random variables in $\mathbb{X}_S$ find one with a minimal distance from $X$. Denote the elements of $S$ in increasing order by $S=\{x_1<\dots<x_m\}$ and let $x_0=-\infty$ and $x_{m+1}=\infty$. Consider the following weight function:

\begin{definition}\label{def:weight} For $0\leq i \leq m$ let
	\[
	w(x_i,x_{i+1})=
	\begin{cases}
	P(x_i < X < x_{i+1}) & \text{if $i=0$ or $i = m$;} \\
	P(x_i < X < x_{i+1})/2 & \text{otherwise.} \\	
	\end{cases}
	\]
\end{definition}
For each $1 <  i \leq m$ let $\hat x_i$ be the maximal element of $\support(X)$ that is smaller than $x_i$. 
Note that, for $1\leq i \leq m$, $P(x_i < X < x_{i+1}) = F_X(\hat x_{i+1}) - F_X(x_i)$, a fact that we will use throughout this section.

\begin{definition}\label{def:error}
Let $\varepsilon(X,S) = \max\limits_{i=0,\dots,m} w(x_{i}, x_{i+1})$.
\end{definition}

We first show that $\varepsilon(X,S)$ is a lower bound for the distance between random variable in $\mathbb{X}_S$ and $X$. Then, we present a random variable $X'\in \mathbb{X}_S$ such that $d_K(X,X')=\varepsilon(X,S)$. It then follows that $X'$ is closets to $X$ in $\mathbb{X}_S$.

%The intuition behind choosing these specific weights and $\varepsilon(X,S)$ being a lower bound is as follows.  Since for every $X'\in\mathbb{X}_S$ the probability mass of $X'$ for the elements not in $S$ are set to $0$, we have that $F_{X'}(\hat x_{i+1})=F_{X'}(x_i)$. Therefore the distance between $X'$ and $X$ at points $x_i$ and $\hat x_{i+1}$ that we have to take into additional account is increased by $F_X(\hat x_{i+1})-F_X(x_i) = P(x_i < X < x_{i+1})$.

%Formally we have the following.

\begin{proposition}\label{prop:minimal}
	If $X'\in\mathbb{X}_S$ then $d_K(X,X') \geq \varepsilon(X,S)$.
\end{proposition}

\begin{proof}

	By definition, for each $0\leq i\leq m$, $d_K(X,X') \geq \max \{|F_X(\hat x_{i+1}) - F_{X'}(\hat x_{i+1})|,|F_X(x_i) - F_{X'}(x_i)| \}$. Note that $F_{X'}(\hat x_{i+1})=F_{X'}(x_i)$ since the probability of  elements not in $S$ is  $0$.

	If $i=0$, that is $x_i=-\infty$, we have that $F_X(x_i)=F_{X'}(x_i)=F_{X'}(\hat x_{i+1})=0$ and therefore $d_K(X,X') \geq |F_X(\hat x_{i+1})| = |F_X(\hat x_{i+1}) - F_{X}(x_i)| =  P(x_i < X < x_{i+1})= w(x_i,x_{i+1})$.

	If $i =m$, that is $x_{i+1}=\infty$, we have that $F_X(\hat x_{i+1})=F_{X'}(\hat x_{i+1})=F_{X'}(x_i)=1$. 
	and therefore $d_K(X,X') \geq |1-F_X(\hat x_i)| = |F_X(\hat x_{i+1}) - F_{X}(x_i)| =  P(x_i < X < x_{i+1}) = w(x_i,x_{i+1})$.

	%  we have that $F_X(x_i)=F_{X'}(x_i)=F_{X'}(\hat x_{i+1})=0$ 
	% 
	% 
	% 
	% We break the proof for the closed intervals where each support variable $x_i$ is a proper number, and for the open intervals where a support variable is either $-\infty$ or $\infty$.

	Otherwise for every $1\leq i< m$,  we use the fact that $max\{|a|,|b|\} \geq |a-b|/2$ for every $a,b\in\mathbb{R}$, to deduce that $d_K(X,X') \geq 1/2| F_X(\hat x_{i+1}) - F_X(x_i) + F_{X'}(x_i) -F_{X'}(\hat x_{i+1})|$. So $d_K(X,X') \geq 1/2| F_X(\hat x_{i+1}) - F_X(x_i) | =P(x_1 < X < x_2)/2 = w(x_i,x_{i+1})$.

Since $d_K(X,X') {\geq}  w(x_i,x_{i+1})$ for any $0{\leq} i{\leq} m$, the proof follows by the definition of $\varepsilon(X,S)$.
\end{proof}

We next describe a random variable $X'\in\mathbb{X}_S$ in a distance of $\varepsilon(X,S)$ from $X$ by its probability mass function:

\begin{definition}\label{def:construction}
	Let $f_{X'}(x_{i}) = w(x_{i-1},x_i) + w(x_i,x_{i+1}) + f_{X}(x_i)$ for $i=1,\dots,m$ and $f_{X'}(x)=0$ for $x \notin S$.
\end{definition}

We first show that $X'$ is a properly defined random variable:

\begin{lemma}
	$f_{X'}$ is a probability mass function. 
\end{lemma}

\begin{proof}
	By definition $f_{X'}(x_{i})\geq 0$ for every $i$. To see that $\sum_i f_{X'}(x_{i}) =1$, 
	we have $\sum_i f_{X'}(x_{i}) = \sum_i (w(x_{i-1},x_i) + w(x_i,x_{i+1}) + f_{X}(x_i)) = 	
	\sum _{x_i\in S} f_{X}(x_i)) + w(x_0,x_1)+ \sum_{0 < i < m} 2 w(x_i,x_{i+1}) + w(x_m,x_{m+1}) = 
	\sum _{x_i\in S} P(X{=}x_i) + P(x_0 {<} X {<} x_1)+ \sum_{0 < i < m} P(x_i < X < x_{i+1}) +P(x_m < X < x_{m+1}) = 1$ since this is the entire support of $X$.	
\end{proof}

Note that, given $S$, $X'$ can be constructed in time linear in the size of the support of $X$.
Its main property, of course, is that the distance between the cumulative distribution functions of $X$ and of $X'$ are bounded by $w(x_i,x_{i+1})$, as follows:
\begin{lemma}\label{lem:balance}
	Let $x\in\support(X)$ and $0\leq i\leq m$ be such that $x_i\leq x \leq x_{i+1}$ then 
    \[
    -w(x_i,x_{i+1})\leq F_X(x)-F_{X'}(x)\leq  w(x_i,x_{i+1}).
    \]
    \end{lemma}

\begin{proof}	
	By induction on $i$.	First see that $F_{X'}(x) = 0$  for every $x_0<x < x_1$ and therefore $F_X(x)-F_{X'}(x) = F_X(x)-0 \leq F_X(\hat x_1)= F_X(\hat x_1)- F_X(x_0) = w(x_0,x_1)$. For $ x = x_1$ we have $F_X(x_1)-F_{X'}(x_1)=  F_X(\hat x_1) +f_{X}(x_1) - (w(x_0,x_1) + w(x_1,x_2) + f_{X}(x_1) = 
	w(x_0,x_1) +f_{X}(x_1) - (w(x_0,x_1) + w(x_1,x_2) + f_{X}(x_1)) = -w(x_1,x_2)$.

	Next assume that $F_X(\hat x_i)-F_{X'}(\hat x_i) = w(x_{i-1},x_i)$.
	Then $F_X(x_i)-F_{X'}(x_i) =  F_X(\hat x_i) +f_{X}(x_i) - (w(x_{i-1},x_i) + w(x_i,x_{i+1}) + f_{X}(x_i)) =  w(x_{i-1},x_i) +f_{X}(x_{i}) - (w(x_{i-1},x_i) + w(x_{i},x_{i+1}) + f_{X}(x_{i})) = -w(x_{i},x_{i+1})$.

	As before we have that for all $x_i< x< x_{i+1}$, we have $F_X(x)-F_{X'}(x) = F_X(x)-F_{X'}(\hat x_{i+1}) \leq  F_X(\hat x_{i+1})-F_{X'}(\hat x_{i+1})$. Then $F_X(\hat x_{i+1})-F_{X'}(\hat x_{i+1}) = (F_X(x_i)+ P(x_i < x <x_{i+1})) - F_{X'}(x_i) = -w(x_{i},x_{i+1}) + 2w(x_{i},x_{i+1}) = w(x_{i},x_{i+1}) $.

	Finally, for $x_m\leq x \leq x_{m+1}$ we have that $F_{X'}(x_m)=1$ therefore $F_X(x_m)-F_{X'}(x_m) = (1- P(x_m<X < x_{m+1})) - 1 =  P(x_m<X < x_{m+1}) = w(x_m,x_{m+1})$, and for every $x_m<x<x_{m+1}$ we have $F_X(x) - F_{X'}(x) < (1- P(x_m<X < x_{m+1})) - 1 <  - P(x_m<X < x_{m+1}))  = -w(x_m,x_{m+1})$ as required.
\end{proof}

>>>>

\DeclareMathOperator{\supp}{support}
\DeclareMathOperator{\support}{support}
\DeclareMathOperator{\Trim}{Trim}
\DeclareMathOperator{\LTrim}{LTrim}
\SetKwFunction{mEqOne}{mEqOne} 
\DeclareMathOperator{\KlmApprox}{Approx}
%\DeclareMathOperator{\OptTrim}{KolmogorovApprox}
\DeclareMathOperator{\OptTrim}{OptTrim}

@@

The case study examined in this paper is the problem of task trees with deadlines~\cite{cohen2015estimating,CohenGW18}. Hierarchical planning is a well-established field in AI~\cite{thomas1988hierarchical,erol1994htn,erol1996complexity}, and is still relevant nowadays~\cite{alford2016hierarchical,xiao2017hierarchical}. A hierarchical plan is a method for representing problems of automated planning in which the dependency among tasks can be given in the form of networks, here we focus on hierarchical plans represented by task trees. The leaves in a task tree are \emph{primitive} actions (or tasks), and the internal nodes are either \emph{sequence} or \emph{parallel} actions. The plans we deal with are of stochastic nature, and the task duration is described as probability distributions in the leaf nodes. We assume that the distributions are independent but {\em not} necessarily identically distributed and that the random variables are discrete and have a finite support.

A sequence node denotes a series of tasks that should be performed consecutively, whereas a parallel node denotes a set of tasks that begin at the same time. A \emph{valid} plan is one that is fulfilled before some given \emph{deadline}, i.e., its \emph{makespan} is less than or equal to the deadline. The objective in this context is to compute the probability that a given plan is valid, or more formally computing $P(X<T)$, where $X$ is a random variable representing the makespan of the plan and $T$ is the deadline. The problem of finding the probability that a task tree satisfies a deadline is known to be NP-hard~\cite{mohring2001scheduling}. In fact, even the problem of summing a set of random variables is NP-hard~\cite{mohring2001scheduling}. This is an example of an explicitly given random variable that we need to estimate deadline meeting probabilities for.

@@

The size of a random variable is measured here by the size of its support, the set of possible outcomes, $|X|=|\support(X)|=|\{x\colon Pr(X=x) \neq 0\}|$. When probability mass functions are maintained as explicit tables, as done in many implementations of statistical software, the size of the support of the variable is proportional to the amount of memory needed to store it and to the complexity of the computations that manipulate it.

@@

In the technical level, the problem we study in this paper is similar to the problem of approximating a set of 2-D points by a step function. The study of this problem was motivated by query optimisation and histogram constructions in database management systems~\cite{applf12,applf13,applf14,applf17,applf18,Fournier2011} and computational geometry~\cite{diaz2001fitting,fournier2008fitting}. There are, however, two technically significant differences between the problem studied in the context of databases and the problem we analyse in this paper. The first difference is that in the context of approximation of random variables, the step function (which is the cumulative distribution function in our context)  must end with a value one, since we are dealing with random variables which sums to one. The second difference is that the first step is not counted because there is no need to put a value in the support of the approximated random variable to generate this first step. These cannot be addressed by adding a constant (two) to $m$ because the first step is always present and because the requirement to end with the value one, restricts the set of eligible step functions. See Section~\ref{sec:alg} for more details.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An algorithm for optimal approximation}\label{sec:algEfficient}
In the scope of this section, let $X$ be a given random variable with a finite support of size $n$, and let  $0<m\leq n$ be a given complexity bound. The section evolves by presenting two algorithms for finding an optimal $m$-approximation of $X$. The first algorithm is $binsearchApprox(X,m)$ (Algorithm~\ref{alg:naive}) which runs in time of $O(n^2log(n))$ and the second one $linApprox(X,m)$ (Algorithm~\ref{alg:linear}) which runs in time of $O(nlog(n))$. We present both $binsearchApprox(X,m)$ and $linApprox(X,m)$ since Algorithm~\ref{alg:linear} is based on Algorithm~\ref{alg:naive} but carry major run-time improvement.

We begin our dissection by presenting a useful fact that it is enough to limit our search to approximations $X'$s such that $\support(X') \subseteq \support(X)$:

Both of the algorithms presented in this section are using the dual problem in order to find the optimal $m$-approximation of $X$. The dual problem is: given a random variable $X$ with $|\support(X)| = n$, and some $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $|\support(X')|\leq m$ where $0< m \leq n$ and $d_K(X,X')$ is minimal.

We do not assume anything about the representation of the random variable $X$ given as an input to the $dual(X,\varepsilon)$ algorithm. However, we transform the representation of every given $X$, to the form of sorted CDF as described in line 1 in the algorithm, which takes $O(nlog(n))$ run-time and has to done only once. The $dual(X,\varepsilon)$ gets as input a random variable $X$ and some $\varepsilon$, the random variable $X$ is represented as in line 1, $X_{CDF} =\{(x_i, c_i)\}_{i=1}^n$ such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$. In lines 3-4 we perform a single pass over $X_{CDF}$ starting from index 1 to find the last index where the cumulative distribution function of $X$ is less than $\varepsilon$, this index is now $f$. In lines 5-6 in the algorithm we start from the last index of $X_{CDF}$ which is $n$, and in decreasing order seek for the last index where the cumulative distribution function of $X$ is less than $1-\varepsilon$, this index is now $l$. The last part of the algorithm is a pass on all elements in $X_{CDF}$ between the indices $f$ and $l$ in order to construct the $m$-approximation random variable of $X$, $X'$. In lines 9-10 we check if the difference between the first element in the current set of indices (which we call $b$) and the last element in the current set of indices (which we call $e+1$) is less that $2\varepsilon$. If yes, we add the pair $\{(x_{b}, (c_b+c_e)/2 - s) \}$ to the set $S$. Otherwise, we add one to the counter of $e$. Note, the reason we need $s$ is to return the random variable $X'$ in a PDF representation since our main motivation is sum of random variables, this part takes $O(n)$ run-time.

@@

\begin{theorem}\label{the:complexityDual}
	The $dual(X,\varepsilon)$ algorithm runs in time $O(n)$, using $O(n)$ memory where $n=|\support(X)|$.
\end{theorem}
\begin{proof}
	This algorithm describes a single pass over the random variable $X$. Lines 3-4 and lines 5-6 are easy to follow, each takes $O(n)$ in the worst case. Lines 8-12 also describe one pass over $X$ since the counter $e$ is updated to $e+1$ at most $n$ times. All together we get run-time complexity of $O(3n)$ = $O(n)$. We are constructing the set $S$ which is of size $n$ in the worst case, therefore, memory complexity is also $O(n)$. 
\end{proof}

The first technique we present is the $binsearchApprox(X,m)$ algorithm which is based on the work done by D{\'i}az-B{\'a}nez and Mesa~\cite{diaz2001fitting}. There are couple of significant changes between our $binsearchApprox(X,m)$ algorithm and the algorithm suggested by D{\'i}az-B{\'a}nez and Mesa~\cite{diaz2001fitting} addressing the differences presented in the Related Work section~\ref{sec:relwork}. The algorithm $binsearchApprox(X,m)$ gets as input a random variable $X$ and some number $m$. Again, we do not mind the original representation of $X$ since we can transform it to a sorted CDF representation in $O(nlog(n))$ run-time as in line 1 which we call $X_{CDF}$. In lines 3-8 we compute all possible errors, in other words, all possible $d_{K}(X,X')$ such that $X'$ is an approximation of $X$ and $\support(X') \subseteq \support(X)$. The error is just the difference between the CDF values of every two elements in $X_{CDF}$. After computing the set $E$, we sort it (line 9) in order to perform a binary search in the last part of the algorithm. In lines 10-19 we perform a binary search over the set $E$, in every step of the binary search we run the $X' = dual(X,\varepsilon)$ algorithm. If the size of $|\support(X')|>m$ then we know that the error $\varepsilon$ is too small and need to extend the search in the right side of $E$, if the size of $|\support(X')|<m$ then we know that the error $\varepsilon$ is too big and need to extend the search in the left side of $E$, otherwise we found the correct $m$. We run the search twice to handle the extreme case where $m \notin \{\forall \varepsilon\in E: |dual(X,\varepsilon)|\}$ and return the smallest $m'$ closest to $m$.

\begin{algorithm}
	\DontPrintSemicolon
	Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$.\;
    $E \gets \emptyset$\; 
    \For{$i \gets 1$ \textbf{to} $n-1$}{
    	\For{$j \gets i+1$ \textbf{to} $n-1$}{
        		\If{$i=1$} {
                	$E \gets E \cup \{ c_j \}$
                }
        		$E \gets E \cup \{ (c_j-c_i)/2 \}$
        }
        $E \gets E \cup \{ 1 - c_i \}$
    }
    Let $e_1<\cdots<e_{n'}$ be such that $E=\{e_1,\dots,e_{n'}\}$ \;

@@

	\caption{$binsearchApprox(X,m)$}   
	\label{alg:naive}
\end{algorithm}

\begin{theorem}\label{the:complexityDual}
	The $binsearchApprox(X,m)$ algorithm runs in time $O(n^2log(n))$, using $O(n^2)$ memory where $n=|\support(X)|$.
\end{theorem}
\begin{proof}
	In the first part of the algorithm, lines 2-8, we construct the set $E$ which takes $O(n^2)$ run-time. In the second part of the algorithm, line 9, we sort the set $E$ which takes $O(n^2log(n^2)) = O(n^2log(n))$  run-time. The third part of the algorithm, lines 10-19, describes a binary sort over the set $E$ where in each step of the sorting we run the $dual$ algorithm, which takes $O(n log(n^2)) = O(n log(n))$ run-time. We run this part twice then we get $O(2n log(n))$. All together, the run-time complexity is $O(n^2+n^2log(n)+2n log(n)) = O(n^2log(n))$ and the memory complexity is for storing the set $E$ which is $O(n^2)$.
\end{proof}

As a step towards $linApprox(X,m)$ algorithm, we present the algorithm $saddlebackApprox(X,m)$ that takes $O(n^2)$ run-time, this technique is very similar to $linApprox(X,m)$ algorithm. The algorithm $saddlebackApprox(X,m)$ gets as input a random variable $X$ and some number $m$. Again, we represent $X$ as a sorted CDF in $O(nlog(n))$ run-time as in line 1 and call it $X_{CDF}$. In line 4 We describe implicitly the sorted matrix $E$, and in lines 3-7 we perform a search over this matrix, in every step of the search we run the $X' = dual(X,\varepsilon)$ algorithm. If the size of $|\support(X')|>m$ then we know that the error $e$ is too small and need to extend the search in the lower side of the matrix, if the size of $|\support(X')|<m$ then we know that the error $e$ is too big and need to extend the search in the right side of the matrix and store this $e$ in $S$. We store in set $S$ all possible candidates for the correct $e$ and in line 8 we get the maximal $m$ with the minimal $e$ value.

\begin{algorithm}
	\DontPrintSemicolon
	Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$.\;
    $i \gets 1$, $j \gets \lceil n/2 \rceil$, $S \gets \emptyset$\;

	\While{$i < \lfloor n/2 \rfloor + 1 \wedge j \geq 2$}
	{
% 		$ e \gets \begin{cases}
% 			c_j   & \text{if } i=\lfloor n/2 \rfloor+1; \\
% 			1-c_i & \text{if } j=\lceil n/2 \rceil; \\
% 			(c_{n+1-j} {-} c_i)/2  & \text{otherwise}
% 		\end{cases}$\;

		$X' \gets dual(X,e_{i,j})$, $m' \gets |\supp(X')|$\;

		\lIf {$m'  > m$}
		{
			$j \gets j - 1$
		}
		\lIf {$m' \leq m$} {
			$i \gets i + 1$, $S \gets S \cup (m',e)$			
		}
	}
	$e \gets \min\{e\colon (m',e) \in \underset{(m',e)\in S}{\operatorname{arg\,max}}\, m'\}$\;
	\Return{$dual(X,e)$}

	\caption{$saddlebackApprox(X,m)$}   
	\label{alg:saddleback}
\end{algorithm}

\begin{algorithm}
		\DontPrintSemicolon
		Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i=Pr(X \leq x_i)$ and $\supp(X)=\{x_1 < \cdots < x_n\}$.\;

		$S \gets  \{ ((1, 1)),(\lceil n/2 \rceil+1,\lceil n/2 \rceil)) \}$, $S' \gets \emptyset$\;

		\While{$S \neq S'$}
		{
			$S' \gets S$\;

			$S \gets \displaystyle \bigcup_{s \in S} split(s)$\;

			$e^- {\gets} median(\{ e_{i_1,j_1} \colon ((i_1,j_1),(i_2,j_2)) \in S \}) $\;

			$X' \gets dual(X,e^-)$, $m' \gets |\supp(X')|$\;

			\If {$m' \leq m$} {
				$S \gets S \setminus \{ ((i_1,j_1),(i_2,j_2)) {\in} S \colon e_{i_1,j_1} {>} e^- \}$ \;
			}

			\Else %\If {$m'  > m$}
			{
				$S \gets S \setminus \{ ((i_1,j_1),(i_2,j_2)) {\in} S \colon e_{i_2,j_2} {\leq} e^- \}$ \;
			}

			$e^+ {\gets} median(\{ e_{i_2,j_2} \colon ((i_1,j_1),(i_2,j_2)) \in S \}) $\;
			$X' \gets dual(X,e^+)$, $m' \gets |\supp(X')|$\;

			\If {$m' \leq m$} {
				$S \gets S \setminus \{ ((i_1,j_1),(i_2,j_2)) {\in} S \colon e_{i_1,j_1} {>} e^+ \}$ \;
			}
			\Else %{$m'  > m$}
			{
				$S \gets S \setminus \{ ((i_1,j_1),(i_2,j_2)) {\in} S \colon e_{i_2,j_2} {\leq} e^+ \}$ \;
			}

		}

		$S'' \gets \{(|dual(X,e_{i,j})|,e_{i,j}) \colon ((i,j),(i,j)) \in S\}$\;

		$e \gets \min\{e\colon (m',e) \in \underset{(m',e)\in S''}{\operatorname{arg\,max}}\, m'\}$\;
		\Return{$dual(X,e)$}

		\caption{$linApprox(X,m)$}   
		\label{alg:linear}
	\end{algorithm}

	\begin{procedure}
		\DontPrintSemicolon
		$j^- \gets \lfloor (j_1+j_2)/2 \rfloor$,
		$j^+ \gets \lceil (j_1+j_2)/2 \rceil$\;
		$i^- \gets \lfloor (i_1+i_2)/2 \rfloor$,
		$i^+ \gets \lceil (i_1+i_2)/2 \rceil$\;

		\Return{$
			\{ ((i_1, j_1), (i^-, j^-)),
			((i_1, j^+), (i^-, j_2)),
			((i^+, j_1), (i_2, j^-)),
			((i^+, j^+), (i_2, j_2))    \}$}

		\caption{{$split(((i_1,j_1),(i_2,j_2)))$}()}
		\label{proc:split}
	\end{procedure}

	Let 
	$$e_{i,j} = \begin{cases}
	c_j   & \text{if } i=\lfloor n/2 \rfloor+1; \\
	1-c_i & \text{if } j=\lceil n/2 \rceil; \\
	(c_{n+1-j} {-} c_i)/2  & \text{otherwise}
	\end{cases}$$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{An algorithm for optimal approximation}\label{sec:alg}
% In the scope of this section, let $X$ be a given random variable with a finite support of size $n$, and let  $0<m\leq n$ be a given complexity bound. The section evolves by developing notations and by collecting facts towards an algorithm for finding an optimal $m$-approximation of $X$.

% The first useful fact is that it is enough to limit our search to approximations $X'$s such that $\support(X') \subseteq \support(X)$:

% \begin{lemma}\label{lem:supContained}
% 	For every random variable $X''$ there is a random variable $X'$ such that $\support(X') \subseteq \support(X)$ and $d_{K}(X,X')\leq d_{K}(X,X'')$.
% \end{lemma}

% \begin{proof}
% 	Let $\{x_1 <\cdots<x_n\} = \support(X)$, and let $x_0 = -\infty, x_{n+1}=\infty$. Consider the random variable $X'$ whose probability mass function is
% 	$f_{X'}(x_i) = P(x_{i-1} < X'' \leq x_i)$ for $i=1,\dots,n-1$,  $f_{X'}(x_n) = P(x_{n-1} < X'')$, and $F_{X'}(x)=0$ if $x\notin \support(X)$.  Since we only "shifted" the probability mass of $X''$ to the support of $X$, we have that $f_{X'}$ is a probability mass function and therefore $X'$ is well defined. 	
% 	By construction, $|F_{X}(x_i)-F_{X'}(x_i)| = |F_{X}(x_i)-F_{X''}(x_i)|$ for every  $0 <  i < n$. For $i=n$ we have $|F_{X}(x_n)-F_{X'}(x_n)| = |1-1|=0$.
% 	Since $|F_{X}(x)-F_{X'}(x)| = |F_{X}(x_i)-F_{X'}(x_i)|$ for every  $0\leq   i  \leq  n$  and $x_i<x<x_{i+1}$, we have that $d_K(X,X')=max_{i}|F_{X}(x_i)-F_{X'}(x_i)|\leq max_{i}|F_{X}(x_i)-F_{X''}(x_i)|\leq d_K(X,X'')$.
% \end{proof}

% %Next, note that every random variable $X''$ with support of size at most $m$ that is contained in $\support(X)$ can be described by first setting the (at most $m$) elements of the support of $X''$; then for every such option, determine $X''$ by setting probability values for the elements in the chosen support of $X'$, and setting $0$ for rest of the elements.

% For a set $S\subseteq \support(X)$, let $\mathbb{X}_S$ denote the set of random variables whose supports are contained in $S$. In Step 1 below, we identify a random variable in $\mathbb{X}_S$  that minimizes the Kolmogorov distance from $X$. We denote the Kolmogorov distance between this variable and $X$ by $\varepsilon(X,S)$. Then, in Step 2, we show how to efficiently find a set $S \subseteq \support(X)$ whose size is smaller or equal to $m$ that minimizes $\varepsilon(X,S)$. Then, in Step 3, an optimal $m$-approximation is constructed by taking a minimal approximation in $\mathbb{X}_S$ where $S$ is the set of size $m$ that that minimizes $\varepsilon(X,S)$.

% \subsection*{Step 1: Finding an $X'$ in $\mathbb{X}_S$ that minimizes $d_K(X,X')$}

% We first fix a set $S\subseteq \support(X)$ and among all the random variables in $\mathbb{X}_S$ find one with a minimal distance from $X$. Denote the elements of $S$ in increasing order by $S=\{x_1<\dots<x_m\}$ and let $x_0=-\infty$ and $x_{m+1}=\infty$. Consider the following weight function:

% \begin{definition}\label{def:weight} For $0\leq i \leq m$ let
% 	\[
% 	w(x_i,x_{i+1})=
% 	\begin{cases}
% 	P(x_i < X < x_{i+1}) & \text{if $i=0$ or $i = m$;} \\
% 	P(x_i < X < x_{i+1})/2 & \text{otherwise.} \\	
% 	\end{cases}
% 	\]
% \end{definition}
% For each $1 <  i \leq m$ let $\hat x_i$ be the maximal element of $\support(X)$ that is smaller than $x_i$. 
% Note that, for $1\leq i \leq m$, $P(x_i < X < x_{i+1}) = F_X(\hat x_{i+1}) - F_X(x_i)$, a fact that we will use throughout this section.

% \begin{definition}\label{def:error}
% Let $\varepsilon(X,S) = \max\limits_{i=0,\dots,m} w(x_{i}, x_{i+1})$.
% \end{definition}

% We first show that $\varepsilon(X,S)$ is a lower bound for the distance between random variable in $\mathbb{X}_S$ and $X$. Then, we present a random variable $X'\in \mathbb{X}_S$ such that $d_K(X,X')=\varepsilon(X,S)$. It then follows that $X'$ is closets to $X$ in $\mathbb{X}_S$.

% %The intuition behind choosing these specific weights and $\varepsilon(X,S)$ being a lower bound is as follows.  Since for every $X'\in\mathbb{X}_S$ the probability mass of $X'$ for the elements not in $S$ are set to $0$, we have that $F_{X'}(\hat x_{i+1})=F_{X'}(x_i)$. Therefore the distance between $X'$ and $X$ at points $x_i$ and $\hat x_{i+1}$ that we have to take into additional account is increased by $F_X(\hat x_{i+1})-F_X(x_i) = P(x_i < X < x_{i+1})$.

% %Formally we have the following.
